---
title: "ThienLab4"
author: "Thien Tam Nguyen"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Lahman)
library(stringi)
```

# Lab report #4 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday (after Thanksgiving) to polish things.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

```{r}

library(rvest)
library(dplyr)
library(stringr)
```

```{r}
url <- "https://www.baseball-reference.com/awards/hof.shtml"

webpage <- read_html(url)

tables <- html_table(webpage, fill = TRUE)

print(paste("Number of tables extracted:", length(tables)))

hof_table <- tables[[1]]

print(paste("Number of columns in hof_table:", ncol(hof_table)))
print("Column names before assignment:")
print(colnames(hof_table))

hof_table <- hof_table %>%
  select(-3)

print(paste("Number of columns after removing the empty column:", ncol(hof_table)))
print("Column names after removing the empty column:")
print(colnames(hof_table))

colnames(hof_table) <- c("Year", "Name", "Voted_By", "Inducted_As", "Votes", "Percent_Votes")

print("Column names after assignment:")
print(colnames(hof_table))

hof_table <- hof_table %>%
  rename(
    yearID = Year,
    playerID = Name,
    votedBy = Voted_By,
    inducted = Inducted_As,
    votes = Votes,
    percent_votes = Percent_Votes
  )

print("Column names after renaming:")
print(colnames(hof_table))

hof_table <- hof_table %>%
  na.omit() %>%
  mutate(
    inducted = ifelse(tolower(inducted) == "yes", "Y", "N"),
    votes = as.numeric(gsub(",", "", votes)),
    percent_votes = as.numeric(gsub("%", "", percent_votes))
  )

print("Sample of cleaned hof_table:")
print(head(hof_table))

write.csv(hof_table, "cleaned_hof_data.csv", row.names = FALSE)

print("Cleaned HOF data has been saved to 'cleaned_hof_data.csv'.")

```



```{r}
write.csv(hof_table, "cleaned_hof_data.csv", row.names = FALSE)
summary(hof_table)

```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Assuming hof_table is your cleaned data frame
# If you need to load the cleaned data from a CSV file:
# hof_table <- read.csv("cleaned_hof_data.csv")

# Convert yearID to a proper Date type (if it's not already)
hof_table$yearID <- as.numeric(hof_table$yearID)

# Group the data by year and inducted status (Y/N)
inductee_counts <- hof_table %>%
  group_by(yearID, inducted) %>%
  summarise(count = n()) %>%
  ungroup()

# Plot the number of inductees by year and inducted status (Y/N)
ggplot(inductee_counts, aes(x = yearID, y = count, fill = inducted)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Hall of Fame Inductees by Year (Y vs N)",
       x = "Year",
       y = "Number of Players",
       fill = "Inducted") +
  scale_fill_manual(values = c("N" = "red", "Y" = "green"))
```


```{r}

hof_data <- Lahman::HallOfFame


head(hof_data)

hof_clean <- hof_data %>%
  filter(yearID >= 2000) %>%  # Example filter for years after 2000
  select(playerID, yearID, inducted) %>%  # Select relevant columns
  mutate(inducted = ifelse(inducted == "Y", "Yes", "No"))  # Clean inducted column

summary(hof_clean)

```
```{r}

hof_clean$playerID <- stri_trim_both(hof_clean$playerID)


special_chars <- hof_clean %>%
  filter(stri_detect_regex(playerID, "[^a-zA-Z0-9]"))
head(special_chars)


ggplot(hof_clean, aes(x = yearID, fill = inducted)) +
  geom_bar(stat = "count", position = "stack") +
  labs(title = "Hall of Fame Inductees by Year", x = "Year", y = "Number of Players")


```

